A new simplified manual tour, with application to exploring classification boundaries in high dimensions

### 2022-07-08 paper/paper.Rmd
application explanation of penguins

In the original work, the method for updating component values, for a 2D projection, was built trackball controls in 3D. A 3D manipulation space is created, as illustrated in Figure \ref{manipspace}, where the controlled variable has full range of motion from -1 to 1. Movements of a cursor are recorded and converted into changes in the values of $V_m$ to change it's values in the displayed 2D projection. Movement could also be constrained to be only in horizontal, vertical, radial or angular motions.

@@

To illustrate the usefulness of the manual tour use the penguins data [@penguins]. We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space.

```{r penguins-scatmat, out.width="100%", width=10, height=10, fig.cap="Scatterplot matrix of the (standardised) penguins data."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
ggscatmat(penguins_data, columns = 2:5, col="species") +
  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  theme(legend.position = "bottom")
```

>>>>

The method for updating component values in @cook_manual_1997 (and utilised in @spyrison_spinifex_2020) are prescribed primarily for a 2D projection, to take advantage of (then) newly developed 3D trackball controls made available for computer gaming. The first step was to construct a 3D manipulation space from a 2D projection, as illustrated in Figure \ref{manipspace}. In this space the coefficient of the controlled variable ranges between -1 and 1. Movements of a cursor are recorded and converted into changes in the values of $V_m$ thus changing the displayed 2D projection. The algorithm also provided constraints to horizontal, vertical, radial or angular motions only.

@@

To illustrate the usefulness of the manual tour we use the 4D penguins data [@penguins]. We will show how classification boundaries can be explored and better understood using a manual tour, on projections and slices through 4D space. Figure \ref(fig:penguins-scatmat) shows a scatterplot matrix of this data. There are four variables (`bl = bill_length_mm, bd = bill_depth_mm, fl = flipper_length_mm, bm = body_mass_g`) measuring the size of the penguins from three species (Adelie, Chinstrap and Gentoo). The scatterplot matrix shows that the three species appear to be likely separable, and that at least the Gentoo can be distinguished from the other two species when `bd` is paired with `fl` or `bm`.

```{r penguins-scatmat, out.width="80%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data. The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
ggscatmat(penguins_data, columns = 2:5, col="species") +
  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  theme(legend.position = "bottom")
```

### 2022-07-08 paper/paper.tex
application explanation of penguins

In the original work, the method for updating component values, for a 2D
projection, was built trackball controls in 3D. A 3D manipulation space
is created, as illustrated in Figure \ref{manipspace}, where the
controlled variable has full range of motion from -1 to 1. Movements of
a cursor are recorded and converted into changes in the values of
\(V_m\) to change it's values in the displayed 2D projection. Movement
could also be constrained to be only in horizontal, vertical, radial or
angular motions.

@@

To illustrate the usefulness of the manual tour use the penguins data
\citep{penguins}. We will show how classification boundaries can be
explored and better understood using a manual tour, on projections and
slices through 4D space.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_files/figure-latex/penguins-scatmat-1} \caption{Scatterplot matrix of the (standardised) penguins data.}\label{fig:penguins-scatmat}
\end{figure}

>>>>

The method for updating component values in \citet{cook_manual_1997}
(and utilised in \citet{spyrison_spinifex_2020}) are prescribed
primarily for a 2D projection, to take advantage of (then) newly
developed 3D trackball controls made available for computer gaming. The
first step was to construct a 3D manipulation space from a 2D
projection, as illustrated in Figure \ref{manipspace}. In this space the
coefficient of the controlled variable ranges between -1 and 1.
Movements of a cursor are recorded and converted into changes in the
values of \(V_m\) thus changing the displayed 2D projection. The
algorithm also provided constraints to horizontal, vertical, radial or
angular motions only.

@@

To illustrate the usefulness of the manual tour we use the 4D penguins
data \citep{penguins}. We will show how classification boundaries can be
explored and better understood using a manual tour, on projections and
slices through 4D space. Figure \ref(fig:penguins-scatmat) shows a
scatterplot matrix of this data. There are four variables
(\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
measuring the size of the penguins from three species (Adelie, Chinstrap
and Gentoo). The scatterplot matrix shows that the three species appear
to be likely separable, and that at least the Gentoo can be
distinguished from the other two species when \texttt{bd} is paired with
\texttt{fl} or \texttt{bm}.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{paper_files/figure-latex/penguins-scatmat-1}

}

\caption{Scatterplot matrix of the (standardised) penguins data. The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
\end{figure}
